% ============================================================
% Appendix 3: Implementation Details
% Covers: Bootstrap Analysis, Bayesian GLM Decision Analysis,
%         and Ethical Reasoning Scoring (AUT-inspired)
% ============================================================

\section*{Appendix 3: Implementation Details}
\addcontentsline{toc}{section}{Appendix 3: Implementation Details}

\subsection*{A3.1 Bootstrap Analysis}

\paragraph{Data loading and preprocessing.}
Nine JSON response matrices are loaded (three models $\times$ three responsibility levels:
\texttt{neutral}, \texttt{not\_responsible}, \texttt{responsible}), each containing
280 records.  The model identity is inferred from the filename (\texttt{doubao},
\texttt{gemini}, \texttt{openai}).  Each record's \texttt{content} field is parsed with a
regular expression to extract the binary decision variable:

\begin{equation}
    y_{ij} = \begin{cases}
        1 & \text{if DECISION contains ``SWITCH''} \\
        0 & \text{if DECISION contains ``NO\_SWITCH'' or ``STAY''} \\
        \text{NaN} & \text{if no valid token is found (excluded)}
    \end{cases}
\end{equation}

Records where $y_{ij}$ cannot be determined are discarded before analysis.

\paragraph{Baseline definition.}
The per-model baseline is the pool of observations satisfying
$\text{responsibility} = \texttt{neutral}$ \emph{and} $\text{role} = \texttt{role\_A}$,
collected across all dilemmas.  All bootstrap contrasts are expressed as deviations from this
anchor, yielding the $\Delta P(\text{SWITCH})$ metric defined in Section~4.

\paragraph{Bootstrap procedure.}
For each cell $(m, s, r)$ --- model $m$, responsibility $s$, role $r$ --- the contrast
$\Delta P = \hat{P}_{m,s,r} - \hat{P}_{m,\texttt{neutral},\texttt{role\_A}}$ is estimated
via non-parametric bootstrap with $B = 500$ iterations (seed 42).  In each iteration $b$:

\begin{enumerate}[noitemsep]
    \item Draw $n_x = |\text{cell}|$ observations \emph{with replacement} from the cell,
          compute $\bar{x}^{(b)}$.
    \item Draw $n_y = |\text{baseline}|$ observations \emph{with replacement} from the
          baseline pool, compute $\bar{y}^{(b)}$.
    \item Record $\Delta^{(b)} = \bar{x}^{(b)} - \bar{y}^{(b)}$.
\end{enumerate}

The 95\% CI is the empirical percentile interval
$[\hat{Q}_{0.025},\,\hat{Q}_{0.975}]$ over the 500 draws.
A cell is declared \textbf{significant} when the CI strictly excludes zero; the direction
indicator is $\uparrow$ (positive effect), $\downarrow$ (negative effect), or
\texttt{ns} (non-significant).  Results are saved to
\texttt{bootstrap\_diff\_vs\_baseline.csv}.

\paragraph{Visualisation.}
For each model, a violin plot is generated over the 21 bootstrap distributions
(7 roles $\times$ 3 responsibility levels).  Violins are grouped into three shaded bands
(Neutral / Not Responsible / Responsible); significant cells are marked with a red star
($*$) above the violin body.  IQR boxes and median dots are superimposed on each violin.

% ============================================================
\subsection*{A3.2 Decision Analysis: Binomial Bayesian Mixed GLM}

\paragraph{Model specification.}
Let $y_i \in \{0,1\}$ denote the binary SWITCH decision for observation $i$.  We fit a
Binomial Bayesian Mixed GLM using \texttt{statsmodels.BinomialBayesMixedGLM}:

\begin{equation}
    \text{logit}\bigl(\Pr(y_i = 1)\bigr)
    = \mathbf{x}_i^{\top}\boldsymbol{\beta} + \mathbf{z}_i^{\top}\mathbf{u},
    \quad \mathbf{u} \sim \mathcal{N}(\mathbf{0},\,\sigma^2_{\text{model}}\mathbf{I})
\end{equation}

\noindent where:
\begin{itemize}[noitemsep]
    \item \textbf{Fixed effects} $\boldsymbol{\beta}$: main effects and all pairwise
          interactions of $\text{responsibility}$ (Treatment baseline:
          \texttt{neutral}) $\times$ $\text{role}$ (Treatment baseline:
          \texttt{role\_A}).
    \item \textbf{Random effect}: model-level random intercept
          $\mathbf{u} \sim \mathcal{N}(\mathbf{0},\,\sigma^2_{\text{model}}\mathbf{I})$,
          fitted via MAP estimation (\texttt{fit\_map}).
\end{itemize}

Posterior means and standard deviations of all parameters are extracted from the MAP
summary.  Approximate 95\% credible intervals are $\hat{\mu} \pm 1.96\,\hat{\sigma}$; a
parameter is considered distinguishable from zero when the interval excludes zero.

\paragraph{Posterior visualisation.}
Posterior distributions are displayed as scaled Gaussian density curves
$\mathcal{N}(\hat{\mu},\,\hat{\sigma}^2)$ arranged in four groups:
\emph{Intercept}, \emph{Responsibility (main)}, \emph{Role (main)}, and
\emph{Resp $\times$ Role (interaction)}.  Log-SD variance parameters are shown separately.
A red dashed horizontal line at zero marks the null for fixed-effect panels.

\paragraph{Residual diagnostics.}
Pearson residuals are computed as
$r_i = (y_i - \hat{p}_i)\,/\,\sqrt{\hat{p}_i(1-\hat{p}_i)}$
and inspected via four diagnostic panels: (i) Pearson residuals vs.\ fitted probabilities,
(ii) residual histogram, (iii) mean raw residual by role, and (iv) mean raw residual by
responsibility.  Cells with $|\bar{r}| > 0.05$ are flagged in red.

An \emph{optimised} model augments the random-effect structure with a dilemma-level random
intercept when a \texttt{dilemma} column is present in the data, reducing the Pearson
$|r|>2$ outlier rate.

% ============================================================
\subsection*{A3.3 Variance Decomposition via Bayesian VB and NUTS Sampling}

\paragraph{Random-effect design.}
To decompose outcome variance across experimental factors, two separate Bayesian
models are fitted, one for each outcome:

\begin{itemize}[noitemsep]
    \item \textbf{Decision} ($y^{\text{dec}} \in \{0,1\}$): Binomial Bayesian Mixed GLM
          with intercept-only fixed effect.
    \item \textbf{Ethical Reasoning} ($y^{\text{rea}} \in \mathbb{Z}^+$): Poisson Bayesian
          Mixed GLM with intercept-only fixed effect, where $y^{\text{rea}}$ is the word
          count of the REASONING field (clipped to $\geq 1$).
\end{itemize}

Both models include four random-effect components specified as indicator matrices:
Role ($\mathbf{Z}_{\text{role}}$), Responsibility ($\mathbf{Z}_{\text{resp}}$),
Model $\times$ Role ($\mathbf{Z}_{m \times r}$), and
Model $\times$ Responsibility ($\mathbf{Z}_{m \times s}$).
Variance parameters are modelled as $\log\sigma$; MAP/VB estimates are obtained via
\texttt{fit\_vb}.

\paragraph{ICC computation.}
The intraclass correlation for component $k$ is:
\begin{equation}
    \text{ICC}_k = \frac{\sigma^2_k}{\sum_{k'}\sigma^2_{k'} + \sigma^2_{\varepsilon}}
    \times 100\%
\end{equation}

where the latent residual variance is:
\begin{equation}
    \sigma^2_{\varepsilon} =
    \begin{cases}
        \pi^2/3 \approx 3.290 & \text{(Bernoulli logistic)} \\[4pt]
        \ln\!\bigl(1 + 1/\bar{y}^{\text{rea}}\bigr) & \text{(Poisson log)}
    \end{cases}
\end{equation}

\paragraph{Monte Carlo 95\% CIs (VB path).}
$N = 50{,}000$ samples are drawn from the VB posterior of each log-SD parameter
$\log\hat{\sigma}_k \sim \mathcal{N}(\hat{\mu}_k,\,\hat{\sigma}^{\text{VB}}_k)$.
The resulting $\sigma^2_k$ samples are converted to ICC percentages; the 2.5th and 97.5th
percentiles form the reported 95\% CI.

\paragraph{NUTS sampling (PyMC path).}
As a cross-validation, a hierarchical PyMC model is fitted with the same four random
effects and HalfNormal($\sigma=1$) priors on the SD parameters.  NUTS is run with
$4 \times 2{,}000$ post-warmup draws (1,000 tuning steps, \texttt{target\_accept}$=0.9$,
seeds 42 and 43 for Decision and Reasoning respectively).  R-hat diagnostics
($\hat{R} < 1.01$ for all $\sigma$ parameters) confirm chain convergence.
ICC tables are derived from the posterior $\sigma^2$ samples in the same way as the
Monte Carlo VB path.

% ============================================================
\subsection*{A3.4 Ethical Reasoning Scoring (AUT-Inspired)}

\paragraph{Metric definition.}
Ethical reasoning richness is quantified as the number of \textbf{valid, semantically
distinct ideas} per response --- inspired by the \emph{fluency} dimension of the
Alternative Uses Task (AUT; Organisciak et al., 2025).  A valid, semantically distinct idea
is defined as:

\begin{quote}
    A self-contained argument, consideration, or ethical claim that is conceptually
    different from all others in the same response.  Mere restatements, paraphrases, or
    elaborations of the same point count as ONE idea.  Purely rhetorical filler
    (\emph{``this is a complex issue''}) does not count.
\end{quote}

\noindent Examples of distinct ideas in trolley-problem reasoning:
\begin{itemize}[noitemsep]
    \item \emph{``The 20\% failure risk changes the expected outcome calculus.''}
          (probabilistic consequence)
    \item \emph{``Actively causing harm is morally worse than allowing harm.''}
          (act/omission distinction)
    \item \emph{``High responsibility requires avoiding direct culpability.''}
          (duty/accountability)
    \item \emph{``Saving more lives respects the equal worth of persons.''}
          (fairness/equal value)
\end{itemize}

\paragraph{LLM-as-judge pipeline.}
Each response's REASONING field is extracted and submitted to \textbf{Claude Opus 4.6}
(\texttt{claude-opus-4-6}) via the Anthropic Messages API with the following
configuration:

\begin{itemize}[noitemsep]
    \item \texttt{max\_tokens}: 256; \texttt{temperature}: 0.7
    \item \texttt{system}: expert evaluator prompt defining the counting rule (above)
    \item \texttt{user}: REASONING text prefaced with the model's binary decision as context
    \item Expected output: a single integer (bare count, no explanation)
\end{itemize}

The judge's response is parsed with a regular expression extracting the first integer
token; records returning no parseable integer are excluded from aggregate statistics
(failure rate $<1\%$).  API calls are serialised with a 1-second inter-request delay to
respect rate limits.

\paragraph{Output format.}
Scored results are written to
\texttt{result/idea\_scores/scored\_\{source\}\_\{timestamp\}.json}, containing:
\texttt{source\_file}, \texttt{model\_judge}, \texttt{metric},
\texttt{scored\_at}, \texttt{mean\_idea\_count}, \texttt{n\_scored},
\texttt{n\_total}, and the full \texttt{records} list with per-record \texttt{idea\_count}.
Across 2,502 valid responses, the overall mean idea count is $\bar{c} = 3.598$
(range $[2.0,\,5.0]$).

\paragraph{Use in variance decomposition.}
Idea-count scores are aggregated per $(m, s, r)$ cell and entered as the dependent
variable in the Poisson Bayesian Mixed GLM described in Appendix~A3.3, enabling a
parallel variance decomposition for reasoning richness alongside the binary decision
outcome.
