\section{Analysis and Discussion}

\subsection{What the Results Imply}

Role framing acts as a latent steering mechanism: assigning an ethical persona shifts a
model's binary moral choice without any explicit verdict instruction, with cross-role
variance reaching up to 52.5\% of the theoretical maximum (Doubao, Dilemma~1).
Susceptibility to this steering is itself a model-level property --- Doubao is the most
redirectable, OpenAI the least --- and the direction of the responsibility effect reverses
across models, indicating that ``accountability'' is not a semantically stable cue.

\subsection{Failure Cases}

\subsubsection{Non-Monotone Responsibility Effect (OpenAI, Role C)}

OpenAI's Deontological Thinker traces the trajectory $\hat{P} = 0.550 \to 0.275 \to
0.925$ across neutral, not-responsible, and responsible conditions --- the only
non-monotone pattern among all 63 cells. By deontological logic, bearing explicit
responsibility should \emph{reinforce} the categorical prohibition on actively harming one
person; instead it nearly eliminates reluctance to switch. The 95\% CI for the responsible
cell $[0.825, 1.000]$ confirms this is not a sampling artefact but a systematic overriding
of the role's normative content by a generic pro-intervention heuristic.

\subsubsection{Persona Reversal Across Dilemma Types (Gemini, Roles E and G)}

Under neutral responsibility, Gemini's Buddhist Thinker and Engineer rank among the two
\emph{lowest}-switching roles in Dilemma~1 ($\hat{P} = 0.075$ and $0.275$). In
Dilemmas~2 and~3 the same prompts produce the two \emph{highest} non-ceiling rates
($\hat{P} = 0.487$ and $0.787$, upward shifts of $+0.412^{*}$ and $+0.512^{*}$). The
rank order of these two personas fully inverts between dilemma types, showing that their
role effects are dilemma-contingent rather than stable trait-level tendencies.

\subsubsection{Directional Disagreement on Christian Framing (Role D)}

The Christian Thinker prompt emphasises compassion, dignity, and stewardship. Under
neutral responsibility it produces $\Delta P = -0.525^{*}$ for Doubao and $-0.225^{*}$
for Gemini (both suppressive), yet $+0.300^{*}$ for OpenAI (elevating switching to
$\hat{P} = 0.975$). The same semantic content resolves to opposite moral verdicts across
models, exposing how model-specific value weighting arbitrates ambiguous ethical vocabulary.

\subsection{Practical Implications}

Role prompting cannot be treated as a model-agnostic alignment tool: a framing that
reliably instils precautionary norms in one model may produce the opposite effect in
another, as demonstrated by the Christian role and the responsibility reversal for OpenAI.
Designers deploying persona-based moral steering in high-stakes applications must validate
the direction and magnitude of each role effect empirically on the target model before
deployment.
