\section{Metrics and Statistical Protocol}

\subsection{Evaluation Metrics}

\subsubsection{Primary Metric: Switch Rate as a Role-Level Score}

The principal evaluation quantity is the \textbf{switch rate} $\hat{P}(\text{SWITCH})$: the
fraction of $n = 40$ repeated trials in which a model issues a SWITCH decision under a given
role--responsibility condition.  In the LLM-ranking framing adopted here, each role assignment
acts as a \emph{system-prompt variant}, and $\hat{P}$ plays the role of a \textbf{win rate}
--- the proportion of trials on which the variant produces the target response class.  The
cell-level score is

\begin{equation}
    \hat{P}_{mrs} = \frac{1}{40}\sum_{j=1}^{40} y_{ij}, \quad y_{ij} \in \{0,1\},
    \quad 1 = \text{SWITCH}
\end{equation}

where $m$ indexes model, $r$ role, and $s$ responsibility level.  Because the three models
carry heterogeneous intrinsic switching tendencies --- Doubao baseline 1.000, OpenAI 0.675,
Gemini 0.250 --- raw $\hat{P}$ values are not directly comparable across models.

\subsubsection{Ranking Contrast Metric: \texorpdfstring{$\Delta P(\text{SWITCH})$}{DeltaP}}

To rank roles \emph{within} each model on a common scale, we define the within-model
deviation

\begin{equation}
    \Delta P_{mrs} = \hat{P}_{mrs} - \hat{P}_{m,\,\text{neutral},\,\text{role\_A}}
\end{equation}

Role\_A (Neutral Analyst, neutral responsibility) serves as the per-model anchor; $\Delta P$
measures how far each condition displaces the model from its unconstrained baseline.  A
positive $\Delta P$ means the role raises switching; negative means it suppresses.  The
observed range is $[-1.00,\,+0.75]$.  Under the neutral responsibility condition, the seven
$\Delta P$ values define a \textbf{role ranking} for each model --- an ordered list of
ethical personas by their tendency to elicit SWITCH decisions.

\subsubsection{Model-Level Ranking Metrics}

While $\hat{P}$ and $\Delta P$ rank \emph{roles within} a model, comparing \emph{models
against each other} requires two complementary model-level scores.

\paragraph{Aggregate Switch Score $\bar{S}_m$.}
The aggregate switch score summarises a model's overall moral tendency across all seven
roles under the neutral responsibility condition:

\begin{equation}
    \bar{S}_m = \frac{1}{7}\sum_{r=1}^{7}\hat{P}_{m,\,r,\,\text{neutral}}
\end{equation}

$\bar{S}_m$ is the model's mean win rate on the task when persona framing is present but
no accountability cue is applied.  A higher $\bar{S}_m$ indicates a stronger intrinsic
pro-switch disposition.  The ranking is: OpenAI ($\bar{S} = 0.761$) $>$ Doubao
($\bar{S} = 0.675$) $>$ Gemini ($\bar{S} = 0.207$).

\paragraph{Steerability Score $\eta_m$.}
The steerability score measures how much the role assignment can redirect the model's
decisions, expressed as the cross-role variance of $\hat{P}$ as a proportion of the
theoretical maximum ($\sigma^2_{\max} = 0.25$):

\begin{equation}
    \eta_m = \frac{1}{0.25}\cdot\frac{1}{7}\sum_{r=1}^{7}
             \bigl(\hat{P}_{m,r,\text{neutral}} - \bar{S}_m\bigr)^{2}
\end{equation}

$\eta_m \in [0,1]$; a value near 1 means role framing produces near-maximal dispersion (half
the roles at 0, half at 1).  The ranking is: Doubao ($\eta = 0.525$) $>$ Gemini
($\eta = 0.300$) $>$ OpenAI ($\eta = 0.123$).

The two dimensions are inversely ordered: OpenAI ranks first on overall switch tendency but
last on steerability, while Doubao ranks first on steerability but second on tendency.  This
trade-off is substantively meaningful --- a high-$\bar{S}$ model has a strong baseline
prior that role framing struggles to override.

\subsubsection{Secondary Metric: Reasoning Richness (Idea Count)}

Beyond the binary decision, each response is assigned an \textbf{idea-count score} by Claude
Opus 4.6 (LLM-as-judge): the number of distinct, non-redundant ethical arguments present in
the free-text reasoning.  Across 2,502 valid responses the distribution has mean 3.598 and
range $[2.0,\,5.0]$, estimated by a linear mixed-effects model (REML intercept $= 3.597$,
SE $= 0.105$).  This metric captures whether a role or responsibility framing changes \emph{how
richly} the model reasons, independently of its binary verdict.

%---------------------------------------------------------------
\subsection{Statistical Analysis}

\subsubsection{Resampling: Non-Parametric Bootstrap}

All uncertainty estimates are obtained by \textbf{non-parametric bootstrap} with $B = 500$
iterations (seed 42).  For a target condition $(m, d, s, r)$, each iteration $b$ draws
$n = 40$ observations with replacement and computes the cell mean $\hat{P}^{(b)}$.  For
model-level aggregate scores, the bootstrap resamples each of the model's cells
independently and averages:

\begin{equation}
    \bar{S}^{(b)}_m = \frac{1}{|C_m|}\sum_{c \in C_m} \hat{P}^{(b)}_c
\end{equation}

where $C_m$ is the set of conditions evaluated for model $m$.  The procedure yields a
bootstrap distribution of 500 values per quantity, from which CIs and reversal
probabilities are derived.

\subsubsection{Uncertainty: Bootstrap Confidence Intervals}

The \textbf{95\% CI} for any quantity is the empirical percentile interval
$[\hat{Q}_{0.025},\,\hat{Q}_{0.975}]$ of the 500 bootstrap draws.  For cell-level scores,
a condition is declared significant when its CI strictly excludes the model's baseline
($\Delta P$ CI excludes zero).  For model-level aggregate scores, two models are declared
reliably ordered when their $\bar{S}^{(b)}$ distributions do not overlap at the 95\% level.

\subsubsection{Ranking Stability}

Model ranking stability is evaluated against two sources of variation treated at equal
weight: \emph{sampling noise} (stochastic draws within a fixed condition) and
\emph{condition variation} (shifting dilemma, responsibility, or role --- none of which is
privileged over the others).  The neutral-responsibility $\times$ neutral-role cell serves
as the reference baseline for computing $\Delta P$ but does not receive special weight in
the ranking assessment.

\paragraph{Rank variance.}
For each bootstrap iteration $b$, model $m$ is assigned an aggregate score
$\bar{S}^{(b)}_m$ by resampling all its condition cells and averaging.  The rank of $m$
across the $B = 500$ iterations defines a rank distribution; its standard deviation
$\text{SD}(\text{rank}_m)$ quantifies how consistently model $m$ occupies a given
position.  A $\text{SD} \approx 0$ indicates a perfectly stable rank; a large SD indicates
susceptibility to rank shifts driven by sampling noise or condition weighting.

\paragraph{Probability of rank reversal.}
For any pair of models $(m_1, m_2)$, the empirical reversal probability is

\begin{equation}
    \Pr\!\bigl[\bar{S}^{(b)}_{m_1} < \bar{S}^{(b)}_{m_2}\bigr]
    = \frac{1}{B}\sum_{b=1}^{B}
      \mathbf{1}\!\bigl[\bar{S}^{(b)}_{m_1} < \bar{S}^{(b)}_{m_2}\bigr]
\end{equation}

A pair is flagged as reversal-susceptible when this probability lies in $(0.05, 0.95)$,
i.e.\ when neither ordering dominates at the 90\% level.  The Gemini rank-3 position is
not reversal-susceptible against either competitor under any condition slice.  The
OpenAI--Doubao rank-1/2 pair is susceptible under the not-responsible slice
($\bar{S}$ gap $= 0.003$) and across Dilemmas~2\&3 where all three models compress into
a 0.044-wide band.

\subsubsection{Variance Decomposition for Reasoning Richness}

To attribute variation in idea-count scores to experimental factors, we fit a fully crossed
\textbf{linear mixed-effects model} (REML) with random effects for model, role,
responsibility, and their pairwise interactions.  The intraclass correlation (ICC), validated
by 200-iteration bootstrap (200/200 converged), is reported in Table~\ref{tab:icc}.

\begin{table}[h]
\centering
\caption{Variance components and ICC for idea-count scores (bootstrap 95\% CIs in brackets).}
\label{tab:icc}
\setlength{\tabcolsep}{6pt}
\begin{tabular}{lrrl}
\hline
\textbf{Component} & $\hat{\sigma}^2$ & \textbf{ICC} & \textbf{95\% CI} \\
\hline
Model ($\alpha_m$)            & 0.0168 & 4.69\%          & $[2.32\%,\;22.87\%]$ \\
Role ($\beta_r$)              & 0.0078 & 2.17\%          & $[0.91\%,\;10.48\%]$ \\
Responsibility ($\gamma_s$)   & 0.0324 & \textbf{9.02\%} & $[2.66\%,\;16.50\%]$ \\
Model $\times$ Role           & 0.0028 & 0.79\%          & $[0.00\%,\;22.41\%]$ \\
Model $\times$ Responsibility & 0.0132 & 3.68\%          & $[0.05\%,\;14.55\%]$ \\
Role $\times$ Responsibility  & 0.0024 & 0.67\%          & $[0.13\%,\;\phantom{0}2.65\%]$ \\
Within ($\varepsilon$)        & 0.2835 & 78.99\%         & $[27.21\%,\;80.96\%]$ \\
\hline
\textbf{Total}                & 0.3589 & 100\%           & --- \\
\hline
\end{tabular}
\end{table}

Responsibility is the dominant between-condition source of reasoning variation
(ICC $= 9.02\%$), while within-condition sampling noise accounts for 79\%, confirming that
$n = 40$ trials per cell is the minimum adequate sample for stable inference on the binary
primary metric.



