\section{Results}

\subsection{Main Results: Model Rankings}

Table~\ref{tab:model_ranking} ranks the three models on two independent scores computed
under the neutral responsibility condition (Dilemma~1, $n=40$ per cell): the aggregate
switch score $\bar{S}_m$ (mean $\hat{P}$ across all seven roles) and the steerability score
$\eta_m$ (cross-role variance as a proportion of the theoretical maximum).  Dilemma, role,
and responsibility are treated as equal-level factors; no dimension is privileged.

\begin{table}[htbp]
\centering
\caption{Model rankings by aggregate switch score $\bar{S}_m$ and steerability $\eta_m$
(neutral, Dilemma~1).}
\label{tab:model_ranking}
\setlength{\tabcolsep}{9pt}
\begin{tabular}{lcccc}
\hline
\textbf{Model} & $\bar{S}_m$ & \textbf{Rank} & $\eta_m$ & \textbf{Rank} \\
\hline
OpenAI & 0.761 & 1 & 0.123 & 3 \\
Doubao & 0.675 & 2 & 0.525 & 1 \\
Gemini & 0.207 & 3 & 0.300 & 2 \\
\hline
\end{tabular}
\end{table}

The two rankings are inversely ordered: OpenAI has the highest baseline switching tendency
but is the least steerable by role framing; Doubao is the most redirectable despite a
lower baseline; Gemini consistently occupies the conservative end on tendency while ranking
second on steerability.

Table~\ref{tab:marginal_means} summarises how these rankings manifest across each factor.

\begin{table}[htbp]
\centering
\caption{Mean $\hat{P}(\text{SWITCH})$ per model across factor levels.  Dilemma and role
columns: neutral responsibility.  Responsibility columns: Dilemma~1, mean over 7 roles.}
\label{tab:marginal_means}
\setlength{\tabcolsep}{3.5pt}
\begin{tabular}{l cc ccc ccccccc}
\hline
& \multicolumn{2}{c}{\textbf{Dilemma}} & \multicolumn{3}{c}{\textbf{Responsibility}}
& \multicolumn{7}{c}{\textbf{Role (neutral, D1)}} \\
\cmidrule(lr){2-3}\cmidrule(lr){4-6}\cmidrule(lr){7-13}
& D1 & D2\&3 & Neu & NR & Resp & A & B & C & D & E & F & G \\
\hline
OpenAI & 0.761 & 0.260 & 0.761 & 0.761 & 0.829
       & 0.675 & 1.000 & 0.550 & 0.975 & 0.700 & 0.550 & 0.875 \\
Doubao & 0.675 & 0.243 & 0.675 & 0.764 & 0.582
       & 1.000 & 1.000 & 0.000 & 0.475 & 0.400 & 0.850 & 1.000 \\
Gemini & 0.207 & 0.216 & 0.207 & 0.296 & 0.186
       & 0.250 & 0.825 & 0.000 & 0.025 & 0.075 & 0.000 & 0.275 \\
\hline
\end{tabular}
\end{table}

%---------------------------------------------------------------
\subsection{Ranking Stability}

\subsubsection{Do Rankings Change Across Conditions?}

\textbf{Dilemma~1.}
Across all three responsibility conditions, Gemini is the only model with a consistent
ordering: it ranks last in every condition ($\bar{S} \in [0.186, 0.296]$), well separated
from the other two.  OpenAI and Doubao, however, swap positions depending on the
responsibility framing.  Under the neutral and not-responsible conditions their scores are
close (0.761 vs.\ 0.761; 0.761 vs.\ 0.764), making the rank-1 assignment statistically
indeterminate.  Under the responsible condition the gap widens sharply (OpenAI 0.829 vs.\
Doubao 0.582), restoring a clear ordering.  Within Dilemma~1, therefore, Gemini is stable
but the top-two ordering is \emph{responsibility-dependent}.

\textbf{Dilemmas~2\&3.}
Shifting to the second and third dilemma types compresses all three models into a narrow
band ($[0.216, 0.260]$, total spread $0.044$).  At this scale the ordering among OpenAI
(0.260), Doubao (0.243), and Gemini (0.216) is consistent in direction but the differences
are too small to be distinguished reliably, rendering the full ranking statistically
indeterminate.  Notably, Gemini's score under Dilemmas~2\&3 (0.216) nearly matches its
Dilemma~1 floor (0.207), whereas OpenAI and Doubao drop substantially from their
Dilemma~1 levels, converging toward Gemini.  The cross-dilemma ordering is therefore
\emph{not stable} for any model pair.

\subsubsection{Are Rankings Stable Across Repeated Trials?}

Within each condition, stochastic sampling (temperature $T = 0.7$, $n = 40$ per cell)
introduces within-condition variance.  The 95\% bootstrap CI widths in
Table~\ref{tab:rank_stability} quantify this at the role level; model-level stability
inherits from role-level stability via the aggregation in $\bar{S}_m$.

\begin{table}[htbp]
\centering
\caption{95\% bootstrap CI width for $\hat{P}$ under neutral responsibility (Dilemma~1).
  Bold = widest cells; 0 = ceiling/floor.}
\label{tab:rank_stability}
\setlength{\tabcolsep}{5pt}
\begin{tabular}{lccccccc}
\hline
\textbf{Model} & A & B & C & D & E & F & G \\
\hline
OpenAI & \textbf{0.300} & 0.000 & \textbf{0.300} & 0.075 & 0.275 & \textbf{0.300} & 0.200 \\
Doubao & 0.000 & 0.000 & 0.000 & \textbf{0.300} & \textbf{0.300} & 0.225 & 0.000 \\
Gemini & 0.275 & 0.225 & 0.000 & 0.075 & 0.175 & 0.000 & 0.275 \\
\hline
\end{tabular}
\end{table}

Doubao's role rankings are highly stable: four of seven roles are degenerate (CI width 0),
and only roles D and E carry uncertainty.  OpenAI has three roles with maximum CI width
(0.300), forming a reversal zone that keeps its aggregate $\bar{S}_m$ imprecise --- the
true rank-1 position relative to Doubao under the not-responsible condition cannot be
resolved at $n = 40$.  Gemini's ranking is stable at the extremes (B first, C/F last) but
uncertain in the middle (A and G are statistically tied).
